---
title: "Correlations Matrix (correlograms) plot for Partial correlations using 'Psych' and 'corrplot  ' package in R"
description: |
  The following code will provide a correlation matrix plot of Partial correlation between multiple variables. 
author:
  - name: Saravana Priyan
date: 03-22-2021
output:
  distill::distill_article:
    self_contained: false
    # code_folding: true
    toc: TRUE
---



Partial correlation is used often to remove the effects of variables like (age, sex, BMI, education) over the correlations of another set of variables.

We will be using the partial.r and corr.p function within Psych package to compute partial correlation matrix.
The function documention can be accessed [here](http://personality-project.org/r/psych/help/partial.r.html).

To create correlogram plot the partial correlation with color coded p values, we will use the library: corrplot in R


### Data Preparation

```{r, include=TRUE, echo = TRUE, size= 'large'}
library(psych)
data <- Tal.Or # data set for example
head(data)
```


### Compute partial correlation 

The partial correlation is computed using the "partial.r" and p values using "corr.p" function in Psych package

```{r, include=TRUE, echo = TRUE, }

#partial gender and age from these relations b/w pmi, import, reaction
par.r <- partial.r(data,1:4,cs(gender,age))
#par.r <- partial.r(data,c("cond", "pmi", "import", "reaction"),cs(gender,age)) # other usage
cp <- corr.p(par.r, n=(nrow(data)-2), adjust="none") #n = nrows - number of covariates
print(cp,short=TRUE)
```

### Plotting using "corrplot" package

Plot the obtained correlation coeff and p value using corrplot with the following code.

```{r, fig.height=7, fig.width = 10}

library(corrplot)

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))


corrplot(par.r,
    method = "color", col = col(200), mar = c(0, 0.5, 2.5, 0),
    addCoef.col = "black", # add correlation coefficient
    number.font = 1, number.cex = 1.5,
    type = "upper", order = "original",
    tl.col = "black", tl.srt = 45, # rotation of text labels
    cl.cex = 1.5,tl.cex = 1.5,
    addgrid.col = "black", bg = "white", 
    title = "correlogram plot the partial correlation",
    # combine with significance level
    p.mat = cp$p , sig.level = 0.05, insig = "blank",
    # hide correlation coefficiens on the diagonal
    diag = FALSE,cl.align.text = 'l'
  )
```

